---
title: "Project2.Rmd"
output: html_document
date: "2022-10-18"
---
Due Oct. 21 at 11:59 PM.

For this first part of the exam, you can either use surveys_complete.csv or your own data. If you are using your own data, you must have data in which you think you have a numerical predictor variable and a numerical response variable. If you are using surveys_complete, you can use weight and hindfoot_length for this.

Save this file in your projects directory. You can either save it in a project two subdirectory, or just put it in projects. Either way is fine.

1. Load in your data. Which variable will you be using as a predictor, and which as a response? (5 pts)

```{r, eval=FALSE}
project2_data <- read_csv("/cloud/project/projects/project_2/eDNA_data.csv")

proj2_data <- project2_data %>% 
  filter(Year!=2001)

projdata2 <- proj2_data %>% rowwise() %>% mutate(n = sum(c_across(Alosa_sp:Trinectes_maculatus)))
View(projdata2)

```
#My predictor variable will be the Water_Temp column and my response variable will be the n column, the number of species per reef location

2. Plot the two against each other with a scatter plot. Do the data appear to be related linearly? (5 pts)
```{r}
ggplot(projdata2, aes(x = Water_Temp, y = n)) +
  geom_point()
```

#This data does not appear to be related linearly

3. Fit the linear model. View the summary. (5 pts)
```{r}
eDNA_model <- lm(n~Water_Temp, data = projdata2)

summary(eDNA_model)
```

4. Does the summary make sense when you compare it to the plot you made? Does our model have good predictive power? Evaluate the residual standard error, intercept, and R-Squared in particular. Would you say your predictor predicts the response? (10 pts)
# This model makes sense based on the plot I made. The plot looked like it had no strong linear relationship and the p-value (0.1347) reflects no significance. The model does not have good predicitive power. The residual standard error is large reflecting the poor predictive power. The intercept means that if the species number was 0 than the temperature would be -83.397. The R-Squared value tells me that temperature explains 5.783% of the species number. Overall, I would say my predictor variable of temperature does not predict my response variable of species number.

5. Plot the model on the graph. Increase the size of the text so it is comfortably readable at 5 feet. Make sure axis labels are readable and not overlapping with one another. (5 pts)
```{r}
ggplot(projdata2, aes(x = Water_Temp, y = n)) +
  geom_point() +
  geom_smooth(method = "lm", color = "navy", size = 0.5, fill = "deeppink") +
  labs(x = "Water Temperature (C)", y = "Number of Species", title = "Water Temperature v Number of Species") +
  annotate("text",x = 25,y = 40, label = "R^2 == 0.0578", parse = T)
```

6.Check the normality of the residuals. Do they look OK? Are we violating assumptions? (5 pts)
```{r}
library(broom)
broom::augment(eDNA_model) -> eDNA_augment
qqnorm(eDNA_augment$.resid)
qqline(eDNA_augment$.resid, col = "blue")
```
# These residuals are far from our line one either side and therefore are violating our assumptions.

Why is normality of residuals important?

#Normality of residuals is important because if our residuals are normal than our model is a good fit. If they are not normal such as what you see above the model is not a good fit.

7. If you are using surveys_complete: Is there interspecific variation in the linear model? Hint: look at our prior work with facet plots. (15 pts)

If you are not using surveys_complete: Do you think there are groupings within your data that may have a separate linear model? Try grouping the data by that variable and redoing the lm. If this would not make sense for your data, you may also attempt to do multiple predictor variables. (15 pts)
```{r}
ggplot(projdata2, aes(x = Water_Temp, y = n)) +
  geom_point() +
  geom_smooth(method = "lm", color = "navy", size = 0.5, fill = "deeppink") +
  facet_wrap(facets = vars(Year)) +
  labs(x = "Water Temperature (C)", y = "Number of Species", title = "Water Temperature v Number of Species")
```

Part Two

In this portion, you are free to use your own data if you have a categorical predictor variable and a response variable. Otherwise, you may use the columns sex and weight in surveys_complete

1. First, plot the data grouped by sex (5 pts)
```{r}
ggplot(projdata2, aes(x = Reef_Name, y = n)) +
  geom_point()
```

2. Try an ANOVA of this data (5 pt)
```{r}
reef_model <- lm(n~Reef_Name, data = projdata2)
aov(reef_model) -> reef_anova
summary(reef_anova)
```

3. How about a linear model? What information does this give you that an ANOVA can’t? (5 pts)
```{r}
reef_model <- lm(n~Reef_Name, data = projdata2)
summary(reef_model)
```
#Linear models give you R-squared values, residual standard errors and highlights the p-value.

4. Plot the lm with the data, with points colored by sex. (10 pts)
```{r}
plt <- projdata2 %>% 
  group_by(Reef_Name) %>% 
  ggplot(aes(x = Reef_Name, y = n, color = Reef_Name)) +
   geom_point() +
  geom_path(aes(group = Reef_or_Control))
print(plt)
```

5. Choose any model we’ve looked at so far, but use two predictor variables. Perform an LM, plot the results, and state if this changes your interpretation of the relationship between variables (10 pts)
# LM Code Here 
```{r}
reef_model_fit <- lm(n~Reef_Name+Water_Temp, data = projdata2)
summary(reef_model_fit)
```

# Plot Code Here 
```{r}
ggplot(projdata2, aes(x = Water_Temp, y = n, color = Reef_Name)) +
  geom_point() +
  geom_smooth(method = "lm") +
  annotate("text", x = 25, y = 40, label = "R^2 == 0.02131", parse = T)
```

# This does not change my interpretation of the relationship between my variables. Both of these predictor variables do not predict my response variable well. The water temperature and the reef only explain about 2% of the variation between number of species per sampling site.

Part Three

1. Add and commit this document (5 pts)
> library(gitcreds)
> gitcreds_set()

? Enter password or token: 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

#In the Git tab (upper right), select the Project2.Rmd and hit Commit. In the pop-up window type what the document is. Hit commit.

2. Push your changes to github (10 pts)
#In the Git Tab select the green up pointing arrow to push. You should now be able to find my Project2.Rmd file on my Github in my project repository.

MS students
My expectation is that you’ll do this with your own data. If any part of this doesn’t make sense with your data, please get in touch ASAP so we can work it out.


